
---
title: "Experiments comparing salt influence on salt growth vs salt mortality"
author: "Anne-Karin Cooke"
output: word_document
---

## Experiments          
This version conducts a comparison experiment on the influence of salinity on the vegetation functions.
Questions we ask:
- Does it lead to significant differences if salinity impacts are only addressed in the biomass growth function or
only in the mortality function or in both. This will be conducted including the virtual saturation which considers both matric and osmotic effects
in opposite to only using a non-salt related general constant.
Further, this shall be investigated for 4 different climates, each consisting of a combination high and low alpha and lambda, respectively.

1. alpha=0.6, lambda=0.1 # dry
2. alpha=1.5, lambda=0.1
3. alpha=1.5, lambda=1.0
4. alpha=0.6, lambda=1.0 # wet

In this version, Sandy Clay Loam was used 
First, a shallow groundwater table was assumed, Z=100 cm.

Experiments :

## Experiment 1: Salt on growth
-for climates 1-4
-with deep gw, Z=1000 cm # NOT YET
-with high gw, Z =100 cm  

## Experiment 2: Salt on mortality
-for climates 1-4
-with deep gw, Z=1000 cm # NOT YET
-with high gw, Z =100 cm 

# Model set up
In this model, salt enters the system through rainfall and from the groundwater table with the same concentration. (just as an example)

##Constants

Soil properties were derived from standard Australian soils in Neurotheta (Minasny and McBratney, 2002, as cited in Shah et al, 2011).

```{r Constants}

# Sandy Clay Loam
    n<-0.367 # porosity
    # more soil variables for evaporation & losses
    # Hydraulic conductivity
    K_s<-52.08*10 # mm/day
    # campbell's b
    b<-6.4069 # neurotheta sandy clay loam
    # van Genuchten parameters
#     avg <- 0.0521
#     nvg <- 1.237
    s_fc<-0.2677/n # Field capacity
    # This is the bubbling pressure
    psi_s_bar<--1.2E-3 #
 
   h1bar =  -psi_s_bar 
   hb = psi_s_bar*-10^5
 

  Z =1000 # [cm] actual groundwater depth 

# parameters describing the soil
  soilpar <- list(b = b, n = n, s_fc = s_fc, K_s = K_s, 
                psi_s_bar = psi_s_bar, h1bar = h1bar, hb = hb)
    #................................................
    # Vegetation 1 (Grass)
    # paspalum secateum F-I and R-I, 2004

    Zr = 400 # soil depth (cm)   Also Table 2...Fernandez-Illescas and Rodriguez-Iturbe...2001


 # parameters describing the root zone   
  vegpar <- list(Zr = Zr)


# parameters describing plant dynamics and salt features

  alpha_i=0.5 #maximum infiltration rate per day, http://www.fao.org/docrep/s8684e/s8684e0a.htm
         
  k=12 # Saco et al, 2013
  W0=0.2 # Saco et al, 2013
  gmax=0.05 # Saco et al, 2013
  k1=5 # Saco et al, 2013
  c=10  # Saco et al, 2013
  f= 0.4  # f is the soil salt leaching efficiency (whether some salt is retained)
  ConcConst = 0.01 # ConcConst is the concentration of the salt in the infiltrating water in g/l
  CM.gw = 0.1 # salt concentration in groundwater
  d=0.24 # fraction of plant mortality
  
  par <- list(alpha_i=alpha_i,k=k, W0=W0, gmax=gmax, k1=k1, c=c, f=f, ConcConst=ConcConst, CM.gw= CM.gw, d=d)
```

## Infiltration function
```{r Infiltration}
Infil <- function(h,P, par){
  
  I=par$alpha_i*h*(P+par$k*par$W0)/(P+par$k)
  
  return (I)
}
```

## Water uptake function
```{r WaterUptake}

WU <- function(M,P,par){ 
# using Svir in here means scaling Svir back to M, easier to do at Svir in balances  
#  WU=par$gmax*((M*(1+Svir))/(((M*(1+Svir))+par$k1)))*P 
  WU=par$gmax*(M/((M+par$k1)))*P 
  
  return(WU)
}
```

##Experiment 2, Salt on Mortality

Salt influence on plant on growth function

```{r Plant growth}

Gr <- function(M,P,par) { 
  
  Gr = par$c*WU(M,P,par)
  
  return(Gr)
}
```

## Plant mortality function WITHOUT salt influence
```{r Mortality function}
Mo <- function(P,par) {
  # needs to be M/Svir because both are "large" numbers
  # you want a number ~1 for multiplication, or <0.1 for addition
  Mo = P*(par$d)
  
  return(Mo)

}
```

## Vertical Flux function
```{r VerticalFlux}
#******************************************
# based on Salvucci, 1993
L_n <- function(M,Z,soilpar,vegpar) {
   Zr <- vegpar$Zr
   hb <- - soilpar$psi_s_bar*10^5 # (mm?)
   soilpar$s_fc <- (Z/hb)^(-1/soilpar$b)
   
   s=M/(soilpar$n*vegpar$Zr)
  
   psi = hb*s^-soilpar$b

   m=2 + 3/soilpar$b#5.64  # in Salvucci's paper it is called n, but I called it m here to not confuse it with porosity

q <-((Z/hb)^(-m)-(psi/hb)^(-m))/(1+(psi/hb)^(-m)+(m-1)*(Z/hb)^(-m))#/(soilpar$n*Zr) # 


# Mass flux
flux <- soilpar$K_s*q


return(flux)
}

```

```{r }
balances <- function(Rain, par, plotit=T,
                      soilpar,
                      vegpar){
   

  
# Storage vectors for the daily steps are initialized.

   M <- rep(0,length(Rain)) # soil moisture [mm]
   h <- rep(0,length(Rain)) # infiltration depth [mm]
   P <- rep(0,length(Rain)) #biomass density []
   CM<- rep(0,length(Rain)) # Salt concentration in soil water in g/L or g/mm
   SmI<- rep(0,length(Rain)) # Salt mass in infiltrating water [g]
   SmM <- rep(0,length(Rain)) # Salt mass in soil water [g]
   In <- rep(0,length(Rain)) # infiltration [mm]

   Svir <- rep(0,length(Rain)) # virtual saturation
   flux<- rep(0,length(Rain)) # drainage and capillary rise flux, according to sign



# Initial values to start the simulation.

   
   M[1] <- 10
   h[1] <- 10 
   P[1] <- 30
   CM[1]<- 0
   Svir[1] <- M[1]



# We decided to split the numerical calculations for the daily into 12 substeps.

   deltat <- 12 # split in 12 increments

  
# Storage vectors for the substeps are initialized.

   M_sub <- rep(0,deltat)
   h_sub <- rep(0,deltat)
   I_sub <- rep(0,deltat)
   #Q_sub <- rep(0,deltat)
   WU_sub <-rep(0,deltat) # Water uptake in mm
   P_sub <- rep(0,deltat) 
   Gr_sub <- rep(0,deltat) # Growth of biomass
   Mo_sub<- rep(0,deltat) # Mortality of biomass
   SmI_sub <- rep(0,deltat) 
   SmM_sub<- rep(0,deltat) 
   CM_sub<- rep(0,deltat) 
  Svir_sub <- rep(0,deltat) # virtual saturation

   flux_sub<-rep(0,deltat) # calculates leakage loss without evaporation loss

   U_salt <-rep(0,deltat)
   L_salt <-rep(0,deltat)


  
   timeincr= 1/deltat
   
  for (t in 2:length(Rain)){
    
    for (tt in 1:(deltat-1)) {
      
      h.old <- ifelse(tt==1,h[t-1],h_sub[tt])
      P.old <- ifelse(tt==1,P[t-1],P_sub[tt])
      M.old <- ifelse(tt==1,M[t-1],M_sub[tt])
      SmI.old <-ifelse(tt==1,SmI[t-1],SmI_sub[tt])
      CM.old <-ifelse(tt==1,CM[t-1],CM_sub[tt])
      Svir.old <-ifelse(tt==1,Svir[t-1],Svir_sub[tt])


# Balance for water depth on soil


h_sub[tt+1] <- h.old + ifelse(tt==1,(10*Rain[t]),0) - Infil(h.old, P.old,par)*timeincr

# Infiltration

      I_sub[tt] <- Infil(h.old, P.old,par)*timeincr

#  1. Update soil moisture with infiltration

      M_sub[tt + 1] <- M.old + I_sub[tt]      
      
# Now do all plant uptake and growth
# water uptake by plants: include infiltration in available water

      WU_sub[tt] <- WU(M=Svir.old,P.old,par)*timeincr 
      
      # growth rate
      Gr_sub[tt] <- Gr(M=Svir.old, P.old,par)*timeincr 
      # Mortality
      Mo_sub[tt]<- Mo(P.old, par)*timeincr
      # calculate plant biomass balance
      P_sub[tt + 1] <- P.old + Gr_sub[tt]- Mo_sub[tt] 
      
      
# re-calculate water balance
# 2. before leaching
      M_sub[tt + 1] <- M.old + I_sub[tt] - WU_sub[tt] #- L_sub[tt] 

# Calculate salt concentration in the soil

 # 3. calculate leaching and capillary rise amount
  flux_sub[tt+1]<-do.call(L_n,list(M=M_sub[tt+1],Z=Z, soilpar=soilpar,vegpar=vegpar))

# 4. final adjust soil moisture for leaching or capillary rise
 
   M_sub[tt + 1] <-  M_sub[tt + 1] + flux_sub[tt+1]*timeincr


# calculate saltbalance

# Salt leaching
    L_salt[tt+1] <- ifelse(flux_sub[tt+1]<0, par$f*CM_sub[tt+1]*flux_sub[tt+1]*timeincr,0) # leaching of salt

# salt uplfow
    U_salt[tt+1] <- ifelse(flux_sub[tt+1]>0, par$CM.gw*flux_sub[tt+1]*timeincr,0) # rise of salt

# salt mass coming in with infiltration
      SmI_sub[tt+1]<- SmI.old + I_sub[tt]*par$ConcConst 

#salt mass in soil
     SmM_sub[tt+1] <- SmI_sub[tt+1] + U_salt[tt+1] - L_salt[tt+1]
# salt concentration in soil
   CM_sub[tt+1]<- (SmM_sub[tt+1]/M_sub[tt+1])*(1/58.44) # 
    
# Virtual saturation (Shah et al., 2012), here in [mm] to be in the same unit as M
    Svir_sub[tt + 1]<-soilpar$n*vegpar$Zr*((soilpar$h1bar*10^-1)^(1/soilpar$b))*
                     ((soilpar$h1bar*10^-1)*(M_sub[tt + 1]/
                       (soilpar$n*vegpar$Zr))^(-soilpar$b)
                       +(3.6*CM_sub[tt + 1]))^(-1/soilpar$b)

    } 
  
# Aggregating the substep results to daily values.

    P[t] = P_sub[deltat]
    M[t] = M_sub[deltat]
    h[t] = h_sub[deltat]
    CM[t] = CM_sub[deltat]
    SmM[t] = SmM_sub[deltat] 
    SmI[t]=SmI_sub[deltat]
    In[t]= sum(I_sub)
    flux[t] = sum(flux_sub)
    Svir[t] = Svir_sub[deltat]

}


# Plotting
 
if (plotit==T) {  
  plot(M, type="l",ylim=c(-10,150),xlim=c(0,time),xlab=("time [d]"), main=paste("Exp 1  lambda=", lambda[j],"alpha=", alpha[i]))
  points(Rain*10, type="h", col="skyblue")
   
    lines(h,type="l", col="blue")
    abline(h=0, col="Gray50",lwd=2,lty=2)

    lines(SmM,type="l", col="red")
    lines(CM,type="l", col="purple")
    lines(P/10,type="l", col="green")
    lines(flux,type="l", col="orange")

#  legend("topright",cex=1, pt.cex=0.4, c("Moisture [mm]","Rainfall [mm]*10","overland flow depth[mm] ","salt mass in soil water [g]", "salt concentration in soil water [g/l]", "Plant biomass density [g/m^2]/10"),
#          col=c("black","skyblue","blue","red","purple","green"),lty=1)
 

}

Out <- data.frame(P=P,M=M,h=h, CM=CM, SmM=SmM, In=In, flux=flux, Svir=Svir)
return(Out)
}
```


## Changing climates
Alpha and lambda are set to low and high values, respecitively.
```{r Variable Climates}
alpha <- c(0.6,1.5) 
lambda <- c(0.1,1)

Store <- list()
sub_store <- list()
time <- 1000
delta <- 0
set.seed(1000)
```

This loop generates rainfall with the distribution properties alpha[i] and lambda[j].
For each alpha[i] and each lambda[j] the soil water balance function is executed and the results stored in a nested list.
```{r run ClimateLoop}
# set the mortality value
# source("Rainfall.R")
Precip <- function(time,alpha,lambda,delta) {
  # generate a vector of times between rainfall events > delta
  f_P<-floor(rexp(time,lambda*exp(-delta/alpha))) # vector of times between rainfall occurrences (equation 4 & 8)
  # generate a binary vector from this (impulse function)
  binary.vec <- unlist(lapply(1:time,function(i) c(rep(0,f_P[i]),1)))
  R <- rexp(length(binary.vec),1/alpha)*binary.vec 
  return(R[1:time])
}

    for (i in 1:length(alpha)) {
  
for (j in 1:length(lambda)) {
      # generate the rainfall
    Rain <- Precip(time,alpha[i],lambda[j],delta)
    Rainlist <- data.frame(Precip(time,alpha[i],lambda[j],delta))
  
    sub_store[[j]] <-data.frame(alpha_o=rep(alpha[i],time),
                                lambda_o=rep(lambda[j],time),
                                balances(Rain,plotit=T, par=par,
                                soilpar, vegpar))
  }
  Store[[i]] <- sub_store
}
```


## Results of Experiment 1


Plotting M and P for different lambdas

```{r}
require(ggplot2)
lambda_sum <- do.call(rbind,Store[][[2]])
lambda_sum$time <- rep(1:time,length(lambda))

pl <- ggplot(lambda_sum,aes(x=time,y=P, colour="P (plant biomass density [gm-2])")) + geom_line()
pl  <- pl + geom_line(aes(x=time,y=M, colour="Moisture"))  
pl <- pl +  facet_wrap(~lambda_o)   #, colour=lambda_o (put this in aes-brackets) 
pl  + ggtitle("Plant biomass P and soilmoisture M for varying lambdas - Experiment 1") +  geom_line(aes(x=time, y=SmM, colour= "S (soil salt mg/L")) + theme(plot.title = element_text(lineheight=.8, face="bold"))
```

Plotting Soil water salt concentration for different alphas and lambdas

```{r}
lambda_sum_all <- do.call(rbind,do.call(rbind,Store))
lambda_sum_all$time <- rep(rep(1:time,length(lambda)),length(alpha))


pa <- ggplot(lambda_sum_all,aes(x=time,y=SmM,col=as.factor(lambda_o))) + geom_line(linetype=1) 
# pa <- pa + + scale_color_gradient(low="blue", high="red")
pa <- pa  + facet_wrap(~alpha_o) + ggtitle("Soil water salt mass [g] for varying alphas and lambdas - Experiment 1") + theme(plot.title = element_text(lineheight=.8, face="bold"))
pa  
```

Storing results to compare and visualize them later on
```{r}
h1P<- lambda_sum$P
h1M <-lambda_sum$M
h1SmM<-lambda_sum$SmM
h1CM<-lambda_sum$CM

``` 

##Experiment 2, Salt on Mortality

Salt influence on plant mortality, not on growth function

```{r Plantgrowth}

Gr <- function(M,P,par) { 
  
  Gr = par$c*WU(M,P,par)
  
  return(Gr)
}
```

## Plant mortality function WITH SALT INFLUENCE, WITH VIRTUAL SATURATION
```{r Mortality}
Mo <- function(P,M,Svir,par) {
  # needs to be M/Svir because both are "large" numbers
  # you want a number ~1 for multiplication, or <0.1 for addition
  Mo = P*(par$d*(M/Svir))
  
  return(Mo)

}
```

## changed balances function, adjusted for the


```{r }
balances <- function(Rain, par, plotit=T,
                      soilpar,
                      vegpar){
   

  
# Storage vectors for the daily steps are initialized.

   M <- rep(0,length(Rain)) # soil moisture [mm]
   h <- rep(0,length(Rain)) # infiltration depth [mm]
   P <- rep(0,length(Rain)) #biomass density []
   CM<- rep(0,length(Rain)) # Salt concentration in soil water in g/L or g/mm
   SmI<- rep(0,length(Rain)) # Salt mass in infiltrating water [g]
   SmM <- rep(0,length(Rain)) # Salt mass in soil water [g]
   In <- rep(0,length(Rain)) # infiltration [mm]

   Svir <- rep(0,length(Rain)) # virtual saturation
   flux<- rep(0,length(Rain)) # drainage and capillary rise flux, according to sign



# Initial values to start the simulation.

   
   M[1] <- 10
   h[1] <- 10 
   P[1] <- 30
   CM[1]<- 0
   Svir[1] <- M[1]



# We decided to split the numerical calculations for the daily into 12 substeps.

   deltat <- 12 # split in 12 increments

  
# Storage vectors for the substeps are initialized.

   M_sub <- rep(0,deltat)
   h_sub <- rep(0,deltat)
   I_sub <- rep(0,deltat)
   #Q_sub <- rep(0,deltat)
   WU_sub <-rep(0,deltat) # Water uptake in mm
   P_sub <- rep(0,deltat) 
   Gr_sub <- rep(0,deltat) # Growth of biomass
   Mo_sub<- rep(0,deltat) # Mortality of biomass
   SmI_sub <- rep(0,deltat) 
   SmM_sub<- rep(0,deltat) 
   CM_sub<- rep(0,deltat) 
  Svir_sub <- rep(0,deltat) # virtual saturation

   flux_sub<-rep(0,deltat) # calculates leakage loss without evaporation loss

   U_salt <-rep(0,deltat)
   L_salt <-rep(0,deltat)


  
   timeincr= 1/deltat
   
  for (t in 2:length(Rain)){
    
    for (tt in 1:(deltat-1)) {
      
      h.old <- ifelse(tt==1,h[t-1],h_sub[tt])
      P.old <- ifelse(tt==1,P[t-1],P_sub[tt])
      M.old <- ifelse(tt==1,M[t-1],M_sub[tt])
      SmI.old <-ifelse(tt==1,SmI[t-1],SmI_sub[tt])
      CM.old <-ifelse(tt==1,CM[t-1],CM_sub[tt])
      Svir.old <-ifelse(tt==1,Svir[t-1],Svir_sub[tt])


# Balance for water depth on soil


h_sub[tt+1] <- h.old + ifelse(tt==1,(10*Rain[t]),0) - Infil(h.old, P.old,par)*timeincr

# Infiltration

      I_sub[tt] <- Infil(h.old, P.old,par)*timeincr

#  1. Update soil moisture with infiltration

      M_sub[tt + 1] <- M.old + I_sub[tt]      
      
# Now do all plant uptake and growth
# water uptake by plants: include infiltration in available water

      WU_sub[tt] <- WU(M=Svir.old,P.old,par)*timeincr 
      
      # growth rate
      Gr_sub[tt] <- Gr(M=M.old, P.old,par)*timeincr 
      # Mortality
      Mo_sub[tt]<- Mo(P.old,M=M.old, Svir=Svir.old, par)*timeincr
      # calculate plant biomass balance
      P_sub[tt + 1] <- P.old + Gr_sub[tt]- Mo_sub[tt] 
      
      
# re-calculate water balance
# 2. before leaching
      M_sub[tt + 1] <- M.old + I_sub[tt] - WU_sub[tt] #- L_sub[tt] 

# Calculate salt concentration in the soil

 # 3. calculate leaching and capillary rise amount
  flux_sub[tt+1]<-do.call(L_n,list(M=M_sub[tt+1],Z=Z,soilpar=soilpar,vegpar=vegpar))

# 4. final adjust soil moisture for leaching or capillary rise
 
   M_sub[tt + 1] <-  M_sub[tt + 1] + flux_sub[tt+1]*timeincr


# calculate saltbalance

# Salt leaching
    L_salt[tt+1] <- ifelse(flux_sub[tt+1]<0, par$f*CM_sub[tt+1]*flux_sub[tt+1]*timeincr,0) # leaching of salt

# salt uplfow
    U_salt[tt+1] <- ifelse(flux_sub[tt+1]>0, par$CM.gw*flux_sub[tt+1]*timeincr,0) # rise of salt

# salt mass coming in with infiltration
      SmI_sub[tt+1]<- SmI.old + I_sub[tt]*par$ConcConst 

#salt mass in soil
     SmM_sub[tt+1] <- SmI_sub[tt+1] + U_salt[tt+1] - L_salt[tt+1]
# salt concentration in soil
   CM_sub[tt+1]<- (SmM_sub[tt+1]/M_sub[tt+1])*(1/58.44) # 
    
# Virtual saturation (Shah et al., 2012), here in [mm] to be in the same unit as M
    Svir_sub[tt + 1]<-soilpar$n*vegpar$Zr*((soilpar$h1bar*10^-1)^(1/soilpar$b))*
                     ((soilpar$h1bar*10^-1)*(M_sub[tt + 1]/
                       (soilpar$n*vegpar$Zr))^(-soilpar$b)
                       +(3.6*CM_sub[tt + 1]))^(-1/soilpar$b)

    } 
  
# Aggregating the substep results to daily values.

    P[t] = P_sub[deltat]
    M[t] = M_sub[deltat]
    h[t] = h_sub[deltat]
    CM[t] = CM_sub[deltat]
    SmM[t] = SmM_sub[deltat] 
    SmI[t]=SmI_sub[deltat]
    In[t]= sum(I_sub)
    flux[t] = sum(flux_sub)
    Svir[t] = Svir_sub[deltat]

}


# Plotting
 
if (plotit==T) {  
  plot(M, type="l",ylim=c(-10,150),xlim=c(0,time),xlab=("time [d]"), main=paste("Exp 2  lambda=", lambda[j],"alpha=", alpha[i]))
  points(Rain*10, type="h", col="skyblue")
   
    lines(h,type="l", col="blue")
    abline(h=0, col="Gray50",lwd=2,lty=2)

    lines(SmM,type="l", col="red")
    lines(CM,type="l", col="purple")
    lines(P/10,type="l", col="green")
    lines(flux,type="l", col="orange")

#  legend("topright",cex=1, pt.cex=0.4, c("Moisture [mm]","Rainfall [mm]*10","overland flow depth[mm] ","salt mass in soil water [g]", "salt concentration in soil water [g/l]", "Plant biomass density [g/m^2]/10"),
#          col=c("black","skyblue","blue","red","purple","green"),lty=1)
 

}

Out <- data.frame(P=P,M=M,h=h, CM=CM, SmM=SmM, In=In, flux=flux, Svir=Svir)
return(Out)
}
```

## Changing climates
Alpha and lambda are set to low and high values, respecitively.
```{r VariableClimates}
alpha <- c(0.6,1.5) 
lambda <- c(0.1,1)

Store <- list()
sub_store <- list()
time <- 1000
delta <- 0
set.seed(1000)
```

This loop generates rainfall with the distribution properties alpha[i] and lambda[j].
For each alpha[i] and each lambda[j] the soil water balance function is executed and the results stored in a nested list.
```{r runClimateLoop}
#source("Rainfall.R")


for (i in 1:length(alpha)) {
  
for (j in 1:length(lambda)) {
    # generate the rainfall
    Rain <- Precip(time,alpha[i],lambda[j],delta)
    Rainlist <- data.frame(Precip(time,alpha[i],lambda[j],delta)) 
    
    sub_store[[j]] <-data.frame(alpha_o=rep(alpha[i],time),
                                lambda_o=rep(lambda[j],time),
                                balances(Rain,plotit=T, par=par,
                                soilpar, vegpar))
  }
  Store[[i]] <- sub_store
}
```


## Results Experiment 2


Plotting M and P for different lambdas

```{r}
require(ggplot2)
lambda_sum <- do.call(rbind,Store[][[2]])
lambda_sum$time <- rep(1:time,length(lambda))

pl <- ggplot(lambda_sum,aes(x=time,y=P, colour="P (plant biomass density [gm-2])")) + geom_line()
pl  <- pl + geom_line(aes(x=time,y=M, colour="Moisture"))  
pl <- pl +  facet_wrap(~lambda_o)   #, colour=lambda_o (put this in aes-brackets) 
pl  + ggtitle("Plant biomass P and soilmoisture M for varying lambdas - Experiment 2") +  geom_line(aes(x=time, y=SmM, colour= "S (soil salt mg/L")) + theme(plot.title = element_text(lineheight=.8, face="bold"))
```

Plotting Soil water salt concentration for different alphas and lambdas

```{r}
lambda_sum_all <- do.call(rbind,do.call(rbind,Store))
lambda_sum_all$time <- rep(rep(1:time,length(lambda)),length(alpha))


pa <- ggplot(lambda_sum_all,aes(x=time,y=SmM,col=as.factor(lambda_o))) + geom_line(linetype=1) 
# pa <- pa + + scale_color_gradient(low="blue", high="red")
pa <- pa  + facet_wrap(~alpha_o) + ggtitle("Soil water salt mass [g] for varying alphas and lambdas - Experiment 2") + theme(plot.title = element_text(lineheight=.8, face="bold"))
pa  
```

```{r}

h2P<- lambda_sum$P
h2M <-lambda_sum$M
h2SmM<-lambda_sum$SmM
h2CM<-lambda_sum$CM

``` 

## Comparing summaries for exp 1 and exp 2
```{r}
summary(h1P) # Plant biomass exp 1
summary(h2P) # Plant biomass exp 2
summary(h1M) # soil moisture exp 1
summary(h2M) # soil moisture exp 2
summary(h1SmM) # soil salt mass exp 1
summary(h2SmM) # soil salt mass exp 2
summary(h1CM) # soil salt concentration exp 1
summary(h2CM) # soil salt concentration exp 2
``` 

## Comparing the distributions for exp 1 and 2
```{r}

par(mfrow = c(1,2))
hist(h1P, breaks=10, main = 'Plant biomass Experiment 1')
hist(h2P, breaks=10, main = 'Plant biomass Experiment 2')
hist(h1M, breaks=10,main="Soil moisture Exp 1")
hist(h2M, breaks=10,main="Soil moisture Exp 2")
hist(h1SmM, breaks=10,main="soil salt mass Exp 1")
hist(h2SmM, breaks=10,main="soil salt mass Exp 2")
hist(h1CM, breaks=10, main="Salt concentration in soil Exp 1")
hist(h2CM, breaks=10, main="Salt concentration in soil Exp 2")

```

## Testing whether distributions for exp 1 and 2 are significantly different from each other
```{r}
wilcox.test(h1M,h2M) # soil moisture
wilcox.test(h1P,h2P) # plant biomass
wilcox.test(h1SmM,h2SmM) # soil salt mass
wilcox.test(h1CM,h2CM) # soil salt concentration

```

## Conclusive remarks

No significant differences between the two approaches
