---
output: word_document
---
### Principal component analysis

#cleaning the data
##occations where plants die as an indicator (goes NA, if model crashes)
```{r}
# for(j in 1:nrow(Store)){
#   if(is.na(Store$Pzero[j])){
#     Store_failure <- rbind(Store_failure, Store[j,])
#     Store<-Store[-j,]
#   }
# }
```

```{r}
# for(j in 1:nrow(Store)){
#   if(is.na(Store$Pzero[j])){
#     Store_failure <- rbind(Store_failure, Store[j,])
#     Store<-Store[-j,]
#   }
# }

# no crashes in Store! :D Store.txt in the repo is a Monte Carlo simulation with Heavy clay, 800 days and 5000 runs
# 
# nrow(Store)
# head(Store)
# ncol(Store)

```
#### LINEAR REGRESSION

Preparing the data
```{r}

mydata <- read.table("EAGELSON heavy medium clay 5000 runs 800 days.txt")
head(mydata)
attach(mydata)




```

```{r}

# for(i in (9:19))
# { 
#   for(j in (1:8) )
#      
#     {       
#        fit <- lm(mydata[,i] ~ mydata[,j], data=mydata) #does a regression for each column in my csv file against my independent variable 'etch'
#        rsq <- summary(fit)$r.squared
#        writelines(paste(rsq,i,"\n"))
#     }
# }


```



 Mean soilmoisture M
```{r}

M_all <- lm(meanM ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(M_all)

#  Scatterplot Matrices from the car Package
# library(car)
# scatterplot.matrix(mydata[9:19], data=mydata,
#     main="Heavy medium clay, 5000 runs, 800 days")


plot(mydata$Zr,mydata$meanM, main= "Root depth and mean soil moisture")
plot(mydata$d,mydata$meanM, main= "Root depth and mean soil moisture")
plot(mydata$c,mydata$meanM, main= "Plan growth factor and mean soil moisture")

plot(mydata$Z,mydata$meanP, main= "Groundwater depth and mean plant biomass")
plot(mydata$d,mydata$meanP, main= "Plant mortality and mean plant biomass")
plot(mydata$c,mydata$meanP, main= "Plant growth factor and mean plant biomass")
plot(mydata$alpha,mydata$meanP, main= "Alpha and mean plant biomass")
plot(mydata$lambda,mydata$meanP, main= "Lambda and mean plant biomass")

plot(mydata$Z,mydata$maxCM, main= "Groundwater depth and maximum soil salt concentration")
plot(mydata$d,mydata$maxCM, main= "Groundwater depth and maximum soil salt concentration")
plot(mydata$ConcConst,mydata$maxCM, main= "Rain salt mass and maximum soil salt concentration")
plot(mydata$c,mydata$maxCM, main= "Plant growth factor and maximum soil salt concentration")
plot(mydata$alpha,mydata$maxCM, main= "Alpha and maximum soil salt concentration")
plot(mydata$lambda,mydata$maxCM, main= "Lambda and maximum soil salt concentration")

# write.table(summary(M_all)$Coefficients, ""

# # Groundwater depth Z
# MZ <- lm(meanM ~ Z, data=mydata)
# summary(MZ)
# 
# # Root depth Zr
# MZr <- lm(meanM ~ Zr, data=mydata)
# summary(MZr)
# 
# # plant mortality d
# Md <- lm(meanM ~ d, data=mydata)
# summary(Md)
# 
# # salt concentration in rain
# MConcConst <- lm(meanM ~ ConcConst, data=mydata)
# summary(MConcConst)
# 
# # salt concentration in groundwater
# MCM.gw <- lm(meanM ~ CM.gw, data=mydata)
# summary(MCM.gw)
# 
# # plant growth factor c
# Mc <- lm(meanM ~ c, data=mydata)
# summary(Mc)
# 
# # alpha
# Malpha <- lm(meanM ~ alpha, data=mydata)
# summary(Malpha)
# 
# # lambda
# Mlambda <- lm(meanM ~ lambda, data=mydata)
# summary(Mlambda)
#  

### comparing the models



```


###Standard deviation soilmoisture M (sdM)

```{r}

sdM_all <- lm(sdM ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(sdM_all)

# Groundwater depth Z
# sdMZ <- lm(sdM ~ Z, data=mydata)
# summary(sdMZ)
# 
# # Root depth Zr
# sdMZr <- lm(sdM ~ Zr, data=mydata)
# summary(sdMZr)
# 
# # plant mortality d
# sdMd <- lm(sdM ~ d, data=mydata)
# summary(sdMd)
# 
# # salt concentration in rain
# sdMConcConst <- lm(sdM ~ ConcConst, data=mydata)
# summary(sdMConcConst)
# 
# # salt concentration in groundwater
# sdMCM.gw <- lm(sdM ~ CM.gw, data=mydata)
# summary(sdMCM.gw)
# 
# # plant growth factor c
# sdMc <- lm(sdM ~ c, data=mydata)
# summary(sdMc)
# 
# # alpha
# sdMalpha <- lm(sdM ~ alpha, data=mydata)
# summary(sdMalpha)
# 
# # lambda
# sdMlambda <- lm(sdM ~ lambda, data=mydata)
# summary(sdMlambda)
#  

### comparing the models




```

Mean soil SALT mass meanSmM


```{r}

SM_all <- lm(meanSmM ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(sdM_all)
# 
# # Groundwater depth Z
# SMZ <- lm(meanSmM ~ Z, data=mydata)
# summary(SMZ)
# 
# # Root depth Zr
# SMZr <- lm(meanSmM ~ Zr, data=mydata)
# summary(SMZr)
# 
# # plant mortality d
# SMd <- lm(meanSmM ~ d, data=mydata)
# summary(SMd)
# 
# # salt concentration in rain
# SMConcConst <- lm(meanSmM ~ ConcConst, data=mydata)
# summary(SMConcConst)
# 
# # salt concentration in groundwater
# SMCM.gw <- lm(meanSmM ~ CM.gw, data=mydata)
# summary(SMCM.gw)
# 
# # plant growth factor c
# SMc <- lm(meanSmM ~ c, data=mydata)
# summary(SMc)
# 
# # alpha
# SMalpha <- lm(meanSmM ~ alpha, data=mydata)
# summary(SMalpha)
# 
# # lambda
# SMlambda <- lm(meanSmM ~ lambda, data=mydata)
# summary(SMlambda)
#  
# 
# ### comparing the models




```

 Standard deviation of Soil Salt mass sdSmM (maybe not too important)

### Mean plant biomass P 

meanP

```{r}
P_all <- lm(meanP ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(P_all)

# # Groundwater depth Z
# PZ <- lm(meanP ~ Z, data=mydata)
# summary(PZ)
# 
# # Root depth Zr
# PZr <- lm(meanP ~ Zr, data=mydata)
# summary(PZr)
# 
# # plant mortality d
# Pd <- lm(meanP ~ d, data=mydata)
# summary(Pd)
# 
# # salt concentration in rain
# PConcConst <- lm(meanP ~ ConcConst, data=mydata)
# summary(PConcConst)
# 
# # salt concentration in groundwater
# PCM.gw <- lm(meanP ~ CM.gw, data=mydata)
# summary(PCM.gw)
# 
# # plant growth factor c
# Pc <- lm(meanP ~ c, data=mydata)
# summary(Pc)
# 
# # alpha
# Palpha <- lm(meanP ~ alpha, data=mydata)
# summary(Palpha)
# 
# # lambda
# Plambda <- lm(meanP ~ lambda, data=mydata)
# summary(Plambda)
#  

### comparing the models



```


### Standard deviation of plant biomass P (sdP)

```{r}
sdP_all <- lm(sdP ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(sdP_all)

# # Groundwater depth Z
# sdPZ <- lm(sdP ~ Z, data=mydata)
# summary(sdPZ)
# 
# # Root depth Zr
# sdPZr <- lm(sdP ~ Zr, data=mydata)
# summary(sdPZr)
# 
# # plant mortality d
# sdPd <- lm(sdP ~ d, data=mydata)
# summary(sdPd)
# 
# # salt concentration in rain
# sdPConcConst <- lm(sdP ~ ConcConst, data=mydata)
# summary(sdPConcConst)
# 
# # salt concentration in groundwater
# sdPCM.gw <- lm(sdP ~ CM.gw, data=mydata)
# summary(sdPCM.gw)
# 
# # plant growth factor c
# sdPc <- lm(sdP ~ c, data=mydata)
# summary(sdPc)
# 
# # alpha
# sdPalpha <- lm(sdP ~ alpha, data=mydata)
# summary(sdPalpha)
# 
# # lambda
# sdPlambda <- lm(sdP ~ lambda, data=mydata)
# summary(sdPlambda)
#  
# 
# ### comparing the models
# 
# 
# ```
# 
# ### mean soil salt concentration (meanCM)
#  
#  
#  
# ```{r}
# CM_all <- lm(meanCM ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
# summary(CM_all)
# 
# # Groundwater depth Z
# CMZ <- lm(meanCM ~ Z, data=mydata)
# summary(CMZ)
# 
# # Root depth Zr
# CMZr <- lm(meanCM ~ Zr, data=mydata)
# summary(CMZr)
# 
# # plant mortality d
# CMd <- lm(meanCM ~ d, data=mydata)
# summary(CMd)
# 
# # salt concentration in rain
# CMConcConst <- lm(meanCM ~ ConcConst, data=mydata)
# summary(CMConcConst)
# 
# # salt concentration in groundwater
# CMCM.gw <- lm(meanCM ~ CM.gw, data=mydata)
# summary(CMCM.gw)
# 
# # plant growth factor c
# CMc <- lm(meanCM ~ c, data=mydata)
# summary(CMc)
# 
# # alpha
# CMalpha <- lm(meanCM ~ alpha, data=mydata)
# summary(CMalpha)
# 
# # lambda
# CMlambda <- lm(meanCM ~ lambda, data=mydata)
# summary(CMlambda)
#  
# 
# ### comparing the models



```
 
Standard deviation of soil salt concentration sdCM, maybe not too important, can be added later
 
Neither minimum of soil salt concentration minCM?? can later be added, easily

### Maximum soil salt concentration  maxCM
```{r}
maxCM_all <- lm(maxCM ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(maxCM_all)

# # Groundwater depth Z
# maxCMZ <- lm(maxCM ~ Z, data=mydata)
# summary(maxCMZ)
# 
# # Root depth Zr
# maxCMZr <- lm(maxCM ~ Zr, data=mydata)
# summary(maxCMZr)
# 
# # plant mortality d
# maxCMd <- lm(maxCM ~ d, data=mydata)
# summary(maxCMd)
# 
# # salt concentration in rain
# maxCMConcConst <- lm(maxCM ~ ConcConst, data=mydata)
# summary(maxCMConcConst)
# 
# # salt concentration in groundwater
# maxCMCM.gw <- lm(maxCM ~ CM.gw, data=mydata)
# summary(maxCMCM.gw)
# 
# # plant growth factor c
# maxCMc <- lm(maxCM ~ c, data=mydata)
# summary(maxCMc)
# 
# # alpha
# maxCMalpha <- lm(maxCM ~ alpha, data=mydata)
# summary(maxCMalpha)
# 
# # lambda
# maxCMlambda <- lm(maxCM ~ lambda, data=mydata)
# summary(maxCMlambda)
#  
# 
# ### comparing the models


```


### Maximum soil salt concentration  maxCM
```{r}
maxCM_all <- lm(maxCM ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(maxCM_all)

# # Groundwater depth Z
# maxCMZ <- lm(maxCM ~ Z, data=mydata)
# summary(maxCMZ)
# 
# # Root depth Zr
# maxCMZr <- lm(maxCM ~ Zr, data=mydata)
# summary(maxCMZr)
# 
# # plant mortality d
# maxCMd <- lm(maxCM ~ d, data=mydata)
# summary(maxCMd)
# 
# # salt concentration in rain
# maxCMConcConst <- lm(maxCM ~ ConcConst, data=mydata)
# summary(maxCMConcConst)
# 
# # salt concentration in groundwater
# maxCMCM.gw <- lm(maxCM ~ CM.gw, data=mydata)
# summary(maxCMCM.gw)
# 
# # plant growth factor c
# maxCMc <- lm(maxCM ~ c, data=mydata)
# summary(maxCMc)
# 
# # alpha
# maxCMalpha <- lm(maxCM ~ alpha, data=mydata)
# summary(maxCMalpha)
# 
# # lambda
# maxCMlambda <- lm(maxCM ~ lambda, data=mydata)
# summary(maxCMlambda)
#  
# 
# ### comparing the models



```


### Cumulative vertical water flux, cum_flux
```{r}
cf_all <- lm(cum_flux ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(cf_all)

# # Groundwater depth Z
# cfZ <- lm(cum_flux ~ Z, data=mydata)
# summary(cfZ)
# 
# # Root depth Zr
# cfZr <- lm(cum_flux ~ Zr, data=mydata)
# summary(cfZr)
# 
# # plant mortality d
# cfd <- lm(cum_flux ~ d, data=mydata)
# summary(cfd)
# 
# # salt concentration in rain
# cfConcConst <- lm(cum_flux ~ ConcConst, data=mydata)
# summary(cfConcConst)
# 
# # salt concentration in groundwater
# cfCM.gw <- lm(cum_flux ~ CM.gw, data=mydata)
# summary(cfCM.gw)
# 
# # plant growth factor c
# cfc <- lm(cum_flux ~ c, data=mydata)
# summary(cfc)
# 
# # alpha
# cfalpha <- lm(cum_flux ~ alpha, data=mydata)
# summary(cfalpha)
# 
# # lambda
# cflambda <- lm(cum_flux ~ lambda, data=mydata)
# summary(cflambda)
#  

### comparing the models



```



### Plant death, P=0, Pzero
```{r}
Pzero_all <- lm(Pzero ~ Z + Zr + d + ConcConst + CM.gw + c + alpha + lambda, data = mydata)
summary(Pzero_all)

# Groundwater depth Z
# PzeroZ <- lm(Pzero ~ Z, data=mydata)
# summary(PzeroZ)
# 
# # Root depth Zr
# PzeroZr <- lm(Pzero ~ Zr, data=mydata)
# summary(PzeroZr)
# 
# # plant mortality d
# Pzerod <- lm(Pzero ~ d, data=mydata)
# summary(Pzerod)
# 
# # salt concentration in rain
# PzeroConcConst <- lm(Pzero ~ ConcConst, data=mydata)
# summary(PzeroConcConst)
# 
# # salt concentration in groundwater
# PzeroCM.gw <- lm(Pzero ~ CM.gw, data=mydata)
# summary(PzeroCM.gw)
# 
# # plant growth factor c
# Pzeroc <- lm(Pzero ~ c, data=mydata)
# summary(Pzeroc)
# 
# # alpha
# Pzeroalpha <- lm(Pzero ~ alpha, data=mydata)
# summary(Pzeroalpha)
# 
# # lambda
# Pzerolambda <- lm(Pzero ~ lambda, data=mydata)
# summary(Pzerolambda)
#  

### comparing the models



```
