
  
---
title: "1D Model"
author: "Anne-Karin Cooke"
output: word_document
---

## Experiments          


1. alpha=0.6, lambda=0.1 # dry
2. alpha=1.5, lambda=0.1
3. alpha=1.5, lambda=1.0
4. alpha=0.6, lambda=1.0 # wet

In this version, Sandy Clay Loam was used.


# Model set up
In this model, salt enters the system through rainfall and from the groundwater table with the same concentration.

##Constants

Soil properties were derived from standard Australian soils in Neurotheta (Minasny and McBratney, 2002, as cited in Shah et al, 2011).

```{r Constants}

 # Sandy Clay Loam
    n<-0.367 # porosity
    # more soil variables for evaporation & losses
    # Hydraulic conductivity
    K_s<-52.08*10 # mm/day
    # campbell's b
    b<-6.4069 # neurotheta sandy clay loam
    # van Genuchten parameters
#     avg <- 0.0521
#     nvg <- 1.237
    s_fc<-0.2677/n # Field capacity
    # This is the bubbling pressure
    psi_s_bar<--1.2E-3 #
 
   h1bar =  -psi_s_bar 
   hb = psi_s_bar*-10^5 # mm
 

  Z =3000 # [mm] actual groundwater depth 

# parameters describing the soil
  soilpar <- list(b = b, n = n, s_fc = s_fc, K_s = K_s, 
                psi_s_bar = psi_s_bar, h1bar = h1bar, hb = hb)
    #................................................
    # Vegetation 1 (Grass)
    # paspalum secateum F-I and R-I, 2004

    Zr = 400 # soil depth (mm)   Check Also Table 2...Fernandez-Illescas and Rodriguez-Iturbe...2001


 # parameters describing the root zone   
  vegpar <- list(Zr = Zr)


# parameters describing plant dynamics and salt features

  alpha_i=1 #maximum infiltration rate per day (K_s and therefore soil type dependency in balances function code)

  k=12 # Saco et al, 2013
  W0=0.2 # Saco et al, 2013
  gmax=0.05 # Saco et al, 2013
  k1=5 # Saco et al, 2013
  c=10  # Saco et al, 2013
  f= 0.8  # f is the soil salt leaching efficiency (whether some salt is retained)
  ConcConst = 0.1 # ConcConst is the concentration of the salt in the infiltrating water in g/l
  CM.gw = 0.1 # salt concentration in groundwater
  d=0.24 # fraction of plant mortality
  
  par <- list(alpha_i=alpha_i,k=k, W0=W0, gmax=gmax, k1=k1, c=c, f=f, ConcConst=ConcConst, CM.gw= CM.gw, d=d)
```

## Infiltration function
```{r Infiltration}
Infil <- function(h,P, par){
  
  I=par$alpha_i*h*(P+par$k*par$W0)/(P+par$k)
  
  return (I)
}
```

## Water uptake function
```{r WaterUptake}

WU <- function(M,P,par){ 
# using Svir in here means scaling Svir back to M, easier to do at Svir in balances  
#  WU=par$gmax*((M*(1+Svir))/(((M*(1+Svir))+par$k1)))*P 
  WU=par$gmax*(M/((M+par$k1)))*P 
  
  return(WU)
}
```


## Vertical Flux function
```{r VerticalFlux}
#******************************************
# based on Salvucci, 1993
L_n <- function(M,Z,soilpar,vegpar) {
   Zr <- vegpar$Zr
   hb <- - soilpar$psi_s_bar*10^5 # (mm?)
   soilpar$s_fc <- (Z/hb)^(-1/soilpar$b)
   
   s=M/(soilpar$n*vegpar$Zr)
  
   psi = hb*s^-soilpar$b
 
   m=2 + 3/soilpar$b#5.64  # in Salvucci's paper it is called n, but I called it m here to not confuse it with porosity


q <-((Z/hb)^(-m)-(psi/hb)^(-m))/(1+(psi/hb)^(-m)+(m-1)*(Z/hb)^(-m))#/(soilpar$n*Zr) # 

# Mass flux

flux <- soilpar$K_s*q

return(flux)
}



```

```{r Precip}
Precip <- function(time,alpha,lambda,delta) {
  # generate a vector of times between rainfall events > delta
  f_P<-floor(rexp(time,lambda*exp(-delta/alpha))) # vector of times between rainfall occurrences (equation 4 & 8)
  # generate a binary vector from this (impulse function)
  binary.vec <- unlist(lapply(1:time,function(i) c(rep(0,f_P[i]),1)))
  R <- rexp(length(binary.vec),1/alpha)*binary.vec 
  return(R[1:time])
}
```


## Plant growth function 
```{r Plantgrowth}

Gr <- function(M,P,par) { 
  
  Gr = par$c*WU(M,P,par)
  
  return(Gr)
}
```

## Plant mortality function 
```{r Mortality}
Mo <- function(P,M,Svir,par) {
  # needs to be M/Svir because both are "large" numbers
  # you want a number ~1 for multiplication, or <0.1 for addition
  Mo = P*(par$d*(M/Svir))
  
  return(Mo)

}
```

##  balances function

```{r }
balances <- function(Rain, par, plotit=T,
                      soilpar,
                      vegpar){
   

  
# Storage vectors for the daily steps are initialized.

   M <- rep(0,length(Rain)) # soil moisture [mm]
   h <- rep(0,length(Rain)) # infiltration depth [mm]
   P <- rep(0,length(Rain)) #biomass density []
   CM<- rep(0,length(Rain)) # Salt concentration in soil water in g/L or g/mm
   SmI<- rep(0,length(Rain)) # Salt mass in infiltrating water [g]
   SmM <- rep(0,length(Rain)) # Salt mass in soil water [g]
   In <- rep(0,length(Rain)) # infiltration [mm]

   Svir <- rep(0,length(Rain)) # virtual saturation
   flux<- rep(0,length(Rain)) # drainage and capillary rise flux, according to sign



# Initial values to start the simulation.

   
   M[1] <- 10
   h[1] <- 10 
   P[1] <- 30
   CM[1]<- 0
   Svir[1] <- M[1]



# We decided to split the numerical calculations for the daily into 12 substeps.

   deltat <- 12 # split in 12 increments

  
# Storage vectors for the substeps are initialized.

   M_sub <- rep(0,deltat)
   h_sub <- rep(0,deltat)
   I_sub <- rep(0,deltat)
   WU_sub <-rep(0,deltat) # Water uptake in mm
   P_sub <- rep(0,deltat) 
   Gr_sub <- rep(0,deltat) # Growth of biomass
   Mo_sub<- rep(0,deltat) # Mortality of biomass
   SmI_sub <- rep(0,deltat) 
   SmM_sub<- rep(0,deltat) 
   CM_sub<- rep(0,deltat) 
  Svir_sub <- rep(0,deltat) # virtual saturation

   flux_sub<-rep(0,deltat) # calculates leakage loss without evaporation loss

   U_salt <-rep(0,deltat) # Salt mass rising
   L_salt <-rep(0,deltat) # salt mass drained


  
   timeincr= 1/deltat
   
  for (t in 2:length(Rain)){
    
    for (tt in 1:(deltat-1)) {
      
      h.old <- ifelse(tt==1,h[t-1],h_sub[tt])
      P.old <- ifelse(tt==1,P[t-1],P_sub[tt])
      M.old <- ifelse(tt==1,M[t-1],M_sub[tt])
      SmI.old <-ifelse(tt==1,SmI[t-1],SmI_sub[tt])
      CM.old <-ifelse(tt==1,CM[t-1],CM_sub[tt])
      Svir.old <-ifelse(tt==1,Svir[t-1],Svir_sub[tt])


# Balance for water depth on soil


h_sub[tt+1] <- h.old + ifelse(tt==1,(10*Rain[t]),0) - Infil(h.old, P.old,par)*timeincr

# Infiltration
ifelse(h_sub[tt+1]<soilpar$K_s, soilpar$alpha_i <- 1, soilpar$alpha_i<- (1-(h_sub[tt+1]-soilpar$K_s)/h_sub[tt+1]))
     
       I_sub[tt] <- Infil(h.old, P.old,par)*timeincr

#  1. Update soil moisture with infiltration

      M_sub[tt + 1] <- M.old + I_sub[tt]      
      
# Now do all plant uptake and growth
# water uptake by plants: include infiltration in available water

      WU_sub[tt] <- WU(M=Svir.old,P.old,par)*timeincr 
      
      # growth rate
      Gr_sub[tt] <- Gr(M=Svir.old, P.old,par)*timeincr 
      # Mortality
      Mo_sub[tt]<- Mo(P.old,M=M.old, Svir=Svir.old, par)*timeincr
      # calculate plant biomass balance
      P_sub[tt + 1] <- P.old + Gr_sub[tt]- Mo_sub[tt] 
      
      
# re-calculate water balance
# 2. before leaching
      M_sub[tt + 1] <- M.old + I_sub[tt] - WU_sub[tt] #- L_sub[tt] 

# Calculate salt concentration in the soil

 # 3. calculate leaching and capillary rise amount
  flux_sub[tt+1]<-do.call(L_n,list(M=M_sub[tt+1],Z=Z,soilpar=soilpar,vegpar=vegpar))

# 4. final adjust soil moisture for leaching or capillary rise
 
   M_sub[tt + 1] <-  M_sub[tt + 1] + flux_sub[tt+1]*timeincr


# calculate saltbalance

# Salt leaching
    L_salt[tt+1] <- ifelse(flux_sub[tt+1]<0, par$f*CM_sub[tt+1]*flux_sub[tt+1]*timeincr,0) # leaching of salt

# salt upflow
    U_salt[tt+1] <- ifelse(flux_sub[tt+1]>0, par$CM.gw*flux_sub[tt+1]*timeincr,0) # rise of salt

# salt mass coming in with infiltration
      SmI_sub[tt+1]<- SmI.old + I_sub[tt]*par$ConcConst 

#salt mass in soil
     SmM_sub[tt+1] <- SmI_sub[tt+1] + U_salt[tt+1] - L_salt[tt+1]
# salt concentration in soil
   CM_sub[tt+1]<- (SmM_sub[tt+1]/M_sub[tt+1])*(1/58.44) # 
    
# Virtual saturation (Shah et al., 2012), here in [mm] to be in the same unit as M
    Svir_sub[tt + 1]<-soilpar$n*vegpar$Zr*((soilpar$h1bar*10^-1)^(1/soilpar$b))*
                     ((soilpar$h1bar*10^-1)*(M_sub[tt + 1]/
                       (soilpar$n*vegpar$Zr))^(-soilpar$b)
                       +(3.6*CM_sub[tt + 1]))^(-1/soilpar$b)

    } 
  
# Aggregating the substep results to daily values.

    P[t] = P_sub[deltat]
    M[t] = M_sub[deltat]
    h[t] = h_sub[deltat]
    CM[t] = CM_sub[deltat]
    SmM[t] = SmM_sub[deltat] 
    SmI[t]=SmI_sub[deltat]
    In[t]= sum(I_sub)
    flux[t] = sum(flux_sub)
    Svir[t] = Svir_sub[deltat]

}


# Plotting
 
if (plotit==T) {  
  plot(M, type="l",ylim=c(-10,100),xlim=c(0,time),xlab=("time [d]"), main=paste(" lambda=", lambda[j],"alpha=", alpha[i]))
  points(Rain*10, type="h", col="skyblue")
   
    lines(h,type="l", col="blue")
    abline(h=0, col="Gray50",lwd=2,lty=2)

    lines(SmM,type="l", col="red")
    lines(CM,type="l", col="purple")
    lines(P/10,type="l", col="green")
    lines(flux,type="l", col="orange")
  
 
#  legend("topright", title="Variables",cex=1, pt.cex=0.2, c("Moisture [mm]","Rainfall [mm]*10","overland flow depth[mm] ","salt mass in soil water [g]", "salt concentration in soil water [g/l]", "Plant biomass density [g/m^2]/10"),
#          col=c("black","skyblue","blue","red","purple","green"),lty=1)
#  
}

Out <- data.frame(P=P,M=M,h=h, CM=CM, SmM=SmM, In=In, flux=flux, Svir=Svir)
return(Out)
}
```

## Changing climates
Alpha and lambda are set to low and high values, respecitively.
```{r VariableClimates}
alpha <- c(0.6,1.5) 
lambda <- c(0.1,1)

Store <- list()
sub_store <- list()
time <- 500
delta <- 0
set.seed(1000)
```

This loop generates rainfall with the distribution properties alpha[i] and lambda[j].
For each alpha[i] and each lambda[j] the soil water balance function is executed and the results stored in a nested list.
```{r runClimateLoop}
#source("Rainfall.R")


for (i in 1:length(alpha)) {
  
for (j in 1:length(lambda)) {
    # generate the rainfall
    Rain <- Precip(time,alpha[i],lambda[j],delta)
    Rainlist <- data.frame(Precip(time,alpha[i],lambda[j],delta)) 
    
    sub_store[[j]] <-data.frame(alpha_o=rep(alpha[i],time),
                                lambda_o=rep(lambda[j],time),
                                balances(Rain,plotit=T, par=par,
                                soilpar, vegpar))
  }
  Store[[i]] <- sub_store
}
```


## Results 


Plotting M and P for different lambdas

```{r}
require(ggplot2)
lambda_sum <- do.call(rbind,Store[][[2]])
lambda_sum$time <- rep(1:time,length(lambda))

pl <- ggplot(lambda_sum,aes(x=time,y=P, colour="P (plant biomass density [gm-2])")) + geom_line()
pl  <- pl + geom_line(aes(x=time,y=M, colour="Moisture"))  
pl <- pl +  facet_wrap(~lambda_o)   
pl  + ggtitle("Plant biomass P and soilmoisture M for varying lambdas") +  geom_line(aes(x=time, y=SmM, colour= "S (soil salt mg/L")) + theme(plot.title = element_text(lineheight=.4))
```

Plotting Soil water salt concentration for different alphas and lambdas

```{r}
lambda_sum_all <- do.call(rbind,do.call(rbind,Store))
lambda_sum_all$time <- rep(rep(1:time,length(lambda)),length(alpha))


pa <- ggplot(lambda_sum_all,aes(x=time,y=SmM,col=as.factor(lambda_o))) + geom_line(linetype=1) 
# pa <- pa + + scale_color_gradient(low="blue", high="red")
pa <- pa  + facet_wrap(~alpha_o) + ggtitle("Soil water salt mass [g] for varying alphas and lambdas") + theme(plot.title = element_text(lineheight=.4))
pa  
```
